---
title: "Data Challenge 2"
author: "Ziwen Zhang"
date: "2022-10-11"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    fig_height: 7
    fig_width: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### load in the required libraries
```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(stringr)
library(lubridate)
library(patchwork)
library(hrbrthemes)
```

## Problem 1
```{r}
Mr_Trash_Wheel<- read_excel("D:/Cornell/Data_Science_I/DataChallenges/A2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Mr. Trash Wheel") ## load in data

wheel1 <- Mr_Trash_Wheel %>%
  clean_names() %>% ##clean the column names
  filter(dumpster != "NA") %>% ## omit rows without dumpster-specific data
  select(-x15, -x16, -x17, -homes_powered) %>% ## remove all columns that are not needed
  mutate(wheel = "Mr. Trash Wheel") ## identifying data from Mr. Trash Wheel

Professor_Trash_Wheel<- read_excel("D:/Cornell/Data_Science_I/DataChallenges/A2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Professor Trash Wheel") ##load in data

wheel2 <- Professor_Trash_Wheel %>%
  clean_names() %>% ## clean the column names
  filter(dumpster != "NA") %>% ##omit rows without dumpster-specific data
  mutate(wheel = "Professor Trash Wheel") ## identifying data from Professor Trash Wheel

Captain_Trash_Wheel <- read_excel("D:/Cornell/Data_Science_I/DataChallenges/A2/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Captain Trash Wheel") ## load in data

wheel3 <- Captain_Trash_Wheel %>%
  clean_names() %>% ## clean the column names
  filter(dumpster != "NA") %>% ## omit rows without dumpster-specific data
  mutate(wheel = "Captain Trash Wheel") ## identifying data from Captain Trash Wheel

all_trash_wheels <- bind_rows(wheel1, wheel2, wheel3) %>%
  pivot_longer(cols = c("plastic_bottles", "polystyrene", "cigarette_butts", "glass_bottles", "grocery_bags", "chip_bags", "sports_balls", "plastic_bags"),
               names_to = "trash_type", 
               values_to = "number") %>% ## bind the three data frames together into a data frame
  mutate(trash_type = str_replace(trash_type, "_", " ")) %>% ## remove "_"
  mutate(trash_type = str_to_title(trash_type)) ## Capitalize

all_trash_wheels_totals_June_2018 <- all_trash_wheels %>%
  filter(month == "June", 
         year == '2018') %>% ## Filtering the data for only June 2018
  group_by(wheel, 
           trash_type) %>% ## group data by wheel and trash_type
  summarise(total_number = sum(number)) ## calculate the total number of each trash item collected by each trash wheel for June 2018

## make a faceted bar plot (by trash type) of the amount of trash (x-axis) collected by each wheel (y-axis): 
all_trash_wheels_totals_June_2018 %>%
  ggplot(aes(x = total_number,
             y = wheel,
             fill = trash_type)) + ##initialize plot object
  geom_bar(stat = "identity") + ## make a bar plot
  facet_wrap(~trash_type, 
             nrow = 3, 
             scale = "free_x")+ ## facet the plot by trash type
  labs(x = "Total number of trash",
       y = "Wheel Type", ## label x and y axis
       title = "Bar plot of total number of each trash item collected by each trash wheel for June 2018", ## add a title
       fill = "Trash Type") ## relabel the legend

```


## Problem 2
```{r}
snp <- read_csv("D:/Cornell/Data_Science_I/DataChallenges/A2/snp.csv")
unemployment <- read_csv("D:/Cornell/Data_Science_I/DataChallenges/A2/unemployment.csv") ## load in data

snp <- snp %>%
  mutate(newdate = mdy(date)) %>% ## convert the data to a date object "newdate"
  mutate(newdate = if_else(year(newdate) >= 2050, 
                           newdate - dyears(100), 
                           newdate)) ## correct data from 1968 and before

snp <- snp %>%
  mutate(year = year(newdate)) %>% ## create a new variable "year"
  mutate(month = month(newdate)) ## create a new variable "month"


unemployment <- unemployment %>%
  pivot_longer(cols = "Jan":"Dec", 
               names_to = "month", 
               values_to = "rate") %>% ## convert data into long format
  mutate(date = paste(month,"01",Year)) %>% ## Create a date column that takes the month and year and indicates the first day of the month
  mutate(date = mdy(date)) ## reformat the data

## make a plot of both the S&P average and the unemployment rate together
ggplot() + ## initialize ggplot object
  geom_line(data = snp, 
            aes(x=newdate, 
                y = close/100), ## make a line for S&P data
            color = "green", ## change color
            lwd = 1) + ## change line width
  geom_line(data = unemployment, 
            aes(x = date, 
                y = rate), ## make a line for unemployment data
            color = "red", ## change color
            lwd = 1) + ## change line width
  ## custom the Y scales
  scale_y_continuous(name = "Unemployment Rate (%)", ## features of the first axis
                     sec.axis = sec_axis(~.*100, 
                                         name="Average S&P Closing Price")) + ## add a second axis and specify its features
  labs(x = "Time", 
       title = "Unemployment Rate and Average Closing Price over time") + 
  theme(axis.title.y.left = element_text(color = "red", 
                                         size = 13), ## change the color of left y axis title
        axis.title.y.right = element_text(color = "green", 
                                          size = 13)) ## change the color of right y axis title
```

## Problem 3
```{r}
## create snp_average data and change it to the same scale of unemployment data
snp_average <- data.frame(
  snp %>%
    group_by(year, 
             month) %>% ## group data by year and month
    summarise(meanclose = mean(close)) %>% ## calculate the mean of close for each month in each year
    mutate(date = paste(month,"01",year)) %>% ## add the first day to each month
    mutate(date = mdy(date))) ## reformat the data

## join the unemployment data with snp_average and filter for data after the start of 2000
full_join(snp_average, unemployment, by = "date") %>% ## join two dataset by date
  filter(year(date) >= 2000) %>% ## filter years later than 2000
  ggplot(aes(x = rate, 
             y = meanclose)) + ## initialize ggplot object
           geom_point(aes(color = year), 
                      size = 2)+ ## make a scatter plot 
  labs(y = 'S & P Closing Price',
       x = 'Unemployment Rate (%)', ## label x and y axis
       title = 'S & P Closing Price versus Unemployment Rate after 2000', ## add a title
       color = "Year") ## color the points with color by year
```

## Problem 4

In problem 2, the line plot shows the trends of unemployment rate and average closing pricing over time. Over the whole period of time, the unemployment rate fluctuated between around 2.5 and 10, whereas the average S&P closing price increased significantly from 1960s to 2000s and then fluctuated around 1250 then improved to above 2000. 
In problem 3, the scatter plot shows the relationship between S&P closing price and unemployment rate after 2000 year. There is a negative relationship between the S&P closing price and the unemployment rate. The S&P closing price decreases whereas the unemployment rate increases from 2000 to around 2005. However, the S&P closing price increases after 2010 but the unemployment decreases after 2010.
